### 2024-11-15 paper/titlepage.tex
version of paper submitted to JSC

\documentclass[12pt]{article}
\usepackage{notes,hyperref}
\renewcommand{\thefootnote}{\alph{footnote}}

\begin{document}
\subsubsection*{Title}
Computing Direct Sum Decompositions
\subsubsection*{Authors} Devlin Mallory\footnote{Basque Center for Applied Mathematics, Plaza Euskadi 5, 48009 Bilbao, Basque Country,  Spain}~and Mahrud Sayrafi\footnote{MPI für Mathematik in den Naturwissenschaften, Inselstraße 22, 04103 Leipzig, Germany}
\subsubsection*{Corresponding author} Mahrud Sayrafi

\end{document}